---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# rwicc

```{r echo=FALSE, results='asis', message=FALSE, warning=FALSE}
# copied from https://github.com/bcallaway11/did/blob/master/README.Rmd
cat(
  badger::badge_github_actions(action = "R-CMD-check.yaml"),
  badger::badge_doi("10.5281/zenodo.6600579", "green"),
  badger::badge_cran_download("rwicc", "grand-total", "blue"),
  badger::badge_cran_download("rwicc", "last-month", "blue"),
  badger::badge_cran_release("rwicc", "blue"),
  badger::badge_devel("d-morrison/rwicc", "blue"),
  badger::badge_cran_checks("rwicc"),
  badger::badge_codecov("d-morrison/rwicc"),
  badger::badge_last_commit("d-morrison/rwicc")
)
```


`rwicc` ("Regression With Interval-Censored Covariates") is an R software package 
implementing an analysis for a regression model involving an interval-censored covariate,
as described in "Regression with Interval-Censored Covariates: Application to Cross-Sectional Incidence
Estimation" by Morrison, Laeyendecker, and Brookmeyer (Biometrics, 2021): https://onlinelibrary.wiley.com/doi/10.1111/biom.13472.

This analysis uses a joint model for the distributions of the outcome of interest and the interval-censored covariate, which is treated as a latent variable; the model parameters are estimated by maximum likelihood using an EM algorithm. 
The submodel used for the distribution of the interval-censored covariate is
somewhat specific to the application of interest (estimation of the mean duration of a biomarker-defined window period for 
cross-sectional incidence estimation), so this package may not be immediately applicable to other problems. We are publishing it
with the goal of making the results in our paper easier to reproduce and with the hope that others might adapt pieces of this code
for their own applications. Please feel free to [contact us](mailto:dmorrison01@ucla.edu) with any questions about the code or the paper!


## Installation

You can install the current released version from [CRAN](https://cran.r-project.org) with:

``` r
install.packages("rwicc")
```

You can install the development version from [GitHub](https://github.com/) with:

``` r
install.packages("devtools")
devtools::install_github("d-morrison/rwicc")
```

## Examples of use

See: https://d-morrison.github.io/rwicc/articles/how-to-use-rwicc.html

## Code of Conduct

Please note that the rwicc project is released with a [Contributor Code of Conduct](https://d-morrison.github.io/rwicc/CODE_OF_CONDUCT.html). By contributing to this project, you agree to abide by its terms.
