---
title: "how-to-use-rwicc"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{how-to-use-rwicc}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r, include = FALSE}

library(knitr)

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This vignette shows how generate a simulated data set and analyze it using the model and estimation algorithm
described in "Regression with Interval-Censored Covariates: Application to Cross-Sectional Incidence
Estimation" by Morrison, Laeyendecker, and Brookmeyer (2021) in Biometrics (https://doi.org/10.1111/biom.13472).

Here, we simulate the data:

```{r}

library(rwicc)

study_cohort_size = 4500
preconversion_interval_length = 365
hazard_alpha = 1
hazard_beta = 0.5
theta_true = c(0.986, -3.88)

set.seed(1)
sim_data = simulate_interval_censoring(
  theta = theta_true,
  "study_cohort_size" = study_cohort_size,
  "preconversion_interval_length" = preconversion_interval_length,
  "hazard_alpha" = hazard_alpha,
  "hazard_beta" = hazard_beta)

# extract the participant-level and observation-level simulated data:
sim_participant_data = sim_data$pt_data
sim_obs_data = sim_data$obs_data
rm(sim_data)

```

Here, we apply our proposed analysis (this takes a couple of minutes; use argument `verbose = TRUE` to print progress messages):

```{r}  
# this call runs the estimation algorithm for the joint modeling approach:
EM_algorithm_outputs = seroconversion_EM_algorithm(
  obs_level_data = sim_obs_data,
  subject_level_data = sim_participant_data,
  bin_width = 7,
  verbose = FALSE)

# extract the estimated mean window period duration from the algorithm's results:
mu_est_EM = EM_algorithm_outputs$Mu

```

Next, we perform alternative analyses using midpoint imputation and uniform imputation:

```{r}

max_iterations_for_glm_fits = 1000 
# bigglm's default maxit = 8, which is not large enough to ensure convergence for this data.
# I don't think it will ever get close to 1000 iterations though; this is effectively Inf.

# midpoint imputation:
{
  
  library(dplyr)
  library(magrittr)
  library(lubridate)
  library(biglm)
  
  sim_participant_data %<>%
    mutate(S_midpoint = L + (R - L) / ddays(2))
  
  sim_obs_data %<>%
    left_join(by = "ID",
              sim_participant_data %<>% dplyr::select(ID, S_midpoint)) %>%
    mutate(T_midpoint = (O - S_midpoint) / ddays(365))
  
  phi_model_est_midpoint =
    bigglm(
      epsilon = tolerance_for_glm_fits,
      maxit = max_iterations_for_glm_fits,
      quiet = TRUE,
      data = sim_obs_data,
      family = binomial(),
      Y ~ T_midpoint
    )
  
  mu_est_midpoint = compute_mu(coef(phi_model_est_midpoint))
  
}

# uniform imputation:
{
  n_imputations = 100 # 10 would be enough
  
  # create a matrix to store the results from each imputed data set
  imputed_coef_ests = matrix(nrow = n_imputations,
                             ncol = 2,
                             dimnames = list(NULL, c("theta0", "theta1")))
  
  for (i in 1:n_imputations)
  {
    
    sim_participant_data %<>%
      mutate(S_imputed =
               L +
               runif(
                 n = n(),
                 min = 0,
                 max = (R - L) / ddays(1)
               ))
    
    sim_obs_data %<>%
      dplyr::select(-any_of("S_imputed")) %>%
      left_join(sim_participant_data %>% dplyr::select(ID, S_imputed),
                by = "ID") %>%
      mutate(T_imputed = (O - S_imputed) / ddays(365))
    
    phi_model_est_imputed =
      bigglm(
        epsilon = tolerance_for_glm_fits,
        maxit = max_iterations_for_glm_fits,
        quiet = TRUE,
        data = sim_obs_data,
        family = binomial(),
        Y ~ T_imputed
      )
    
    imputed_coef_ests[i, ] = coef(phi_model_est_imputed)
    
  }
  
  imputed_coef_ests_mean = colMeans(imputed_coef_ests)
  
  mu_est_imputed = compute_mu(imputed_coef_ests_mean)
  
}
```

Now, let's graph the results. First, let's plot the true and estimated CDFs for the distribution of seroconversion date,
for individuals who enroll on the first calendar day of the cohort study:


```{r}

library(ggplot2)
cum_haz_fn0 = function(years_since_study_start)
{
  hazard_alpha * years_since_study_start + hazard_beta / 2 * years_since_study_start^2
}

cum_haz_fn = function(years_since_study_start,
                      `years from study start to enrollment`)
{
  cum_haz_fn0(years_since_study_start) - cum_haz_fn0(`years from study start to enrollment`)
}

surv_fn = function(years_since_study_start,
                   `years from study start to enrollment`)
  exp(-cum_haz_fn(
    years_since_study_start,
    `years from study start to enrollment`
  ))

true_model_label = "Data-generating model"
est_model_label = "Estimated model"

lwd1 = 1
omega.hat = EM_algorithm_outputs$Omega %<>%
  mutate(
    "P(S>s|E=0)" = cumprod(`P(S>s|S>=s,E=e)`),
    "P(S>=s|E=0)" = lag(`P(S>s|E=0)`, default = 1) 
  )

plot1 = ggplot(
  aes(
    y = 1 - `P(S>=s|E=0)`,
    x = (S - ymd("2001-01-01")) / ddays(365)
  ),
  data = omega.hat %>% filter(S < max(S))) +
  
  geom_step(
    direction = "hv",
    aes(col = est_model_label,
        linetype = est_model_label),
    lwd = lwd1
  ) +
  
  xlab("Calendar time (years since start of study)") +
  ylab("Probability of seroconversion") +
  geom_function(
    fun = function(x)
      1 - surv_fn(x, 0),
    aes(col = true_model_label,
        linetype = true_model_label),
    lwd = lwd1
  ) +
  
  scale_colour_discrete("") +
  scale_linetype_discrete("") +
  theme(
    axis.title.x = element_text(size = 20),
    axis.text.x = element_text(size = 14),
    axis.title.y = element_text(size = 20),
    axis.text.y = element_text(size = 14),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_blank(),
    axis.line = element_line(colour = "black"),
    legend.position = "top",
    legend.text =
      element_text(size = 14,
                   margin = margin(r = 10, unit = "pt"))
  )

print(plot1)

```

We can see that our joint modeling approach does not always estimate this distribution very accurately at this sample size.
Nevertheless, the next graph will show us that the joint model very accurately estimates the true distribution for $P(Y|T)$
and the true value of $\mu$:

```{r}

library(scales)


mu_true = compute_mu(theta_true)
phi_true = build_phi_function_from_coefs(theta_true)
phi_est_joint_modeling = build_phi_function_from_coefs(EM_algorithm_outputs$Theta)
phi_est_midpoint = build_phi_function_from_coefs(coef(phi_model_est_midpoint))
phi_est_imputed = build_phi_function_from_coefs(imputed_coef_ests_mean)

true_model_label2 = paste(sep = "",
                          "'   data-generating model;'~mu == '",
                          format(mu_true, digits = 1, nsmall = 1),
                          "'~days"
)

midpoint_label = paste(sep = "",
                       "` midpoint imputation;`~hat(mu) == '",
                       format(mu_est_midpoint, digits = 1, nsmall = 1),
                       "'~days"
)

uniform_label = paste(sep = "",
                      "`uniform imputation;`~hat(mu) == '",
                      format(mu_est_imputed, digits = 1, nsmall = 1),
                      "'~days"
)

joint_modeling_label = paste(sep = "",
                             "'  joint modeling;'~hat(mu) =='",
                             format(mu_est_EM, digits = 1, nsmall = 1),
                             "'~days"
)

lwd1 = 1


  plot2 = ggplot(data = NULL) +
    geom_function(
      fun = phi_true,
      lwd = lwd1,
      aes(colour = true_model_label2, linetype = true_model_label2)
    ) +
    geom_function(
      fun = phi_est_midpoint,
      lwd = lwd1,
      aes(colour = midpoint_label, linetype = midpoint_label)
    ) +
    geom_function(
      fun = phi_est_imputed,
      lwd = lwd1,
      aes(colour = uniform_label, linetype = uniform_label)
    ) +
    geom_function(
      fun = phi_est_joint_modeling,
      lwd = lwd1,
      aes(colour = joint_modeling_label, linetype = joint_modeling_label)
    ) +
    xlim(0, 2) +
    ylab("Probability of an MAA-positive blood sample") +
    xlab("Time since seroconversion (years)") +
    scale_colour_discrete("",   labels = label_parse()) +
    scale_linetype_discrete("", labels = label_parse()) +
    theme(
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      panel.background = element_blank(),
      axis.line = element_line(colour = "black")
    ) +
    theme(
      axis.title.x = element_text(size = 20),
      axis.text.x = element_text(size = 14),
      axis.title.y = element_text(size = 18),
      axis.text.y = element_text(size = 14),
      legend.text.align = 1,
      legend.box.just = "right",
      legend.text = element_text(size = 16),
      legend.position = c(.65, .85),
      legend.title = element_blank()
    )


print(plot2)


```

